### 2014-07-02 uTAS2014/paper.tex
started the manuscript

\documentclass{article}
usepackage[utf8]{inputenc}

\title{3D Printed Microfluidic Devices for Oxygen Control in Cell Culture}

\usepackage{natbib}
\usepackage{graphicx}

\usepackage{fontspec}
\setmainfont{Times New Roman}

\begin{document}

\maketitle

\section{INTRODUCTIOM}

This is a reference \cite{Au2014}.

\section{EXPERIMENTAL}

\section{RESULTS AND DISCUSSION}

\begin{figure}[h!]
\centering
\includegraphics[scale=1.7]{3-inch-well-photo.png}
\caption{3" petri dish device}
\label{fig:3-inch-well-photo}
\end{figure}

\section{CONCLUSION}

\bibliographystyle{plain}
\bibliography{references}
\end{document}